---
title: "MacroStatsGraphsMARKDOWN"
author: "Teddy Jorgensen"
date: "2024-08-07"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PART ONE ASSEMBLING THE DATA

Load in all relevant and required packages

```{r}
library(tidyverse)
library(ggplot2)
library(sqldf)
library(ggpubr)
library(broom)
library(AICcmodavg)
```

Read in the macroinvertebrates datasheet from the local computer, and
add in a singular "1" value for every individual macro(this makes
creating some graphs much easier), then upload the land use data

```{r}

MacrosDraft<-read.csv("/Users/tjorgensen/Desktop/Underwood 2024/DATA - Macros.csv")
MacrosDraft$GraphCount<-rep(1,times=count(MacrosDraft))

SiteData<-read.csv("/Users/tjorgensen/Desktop/Underwood 2024/DATA - SiteData.csv")

J3LandUse<-read.csv("/Users/tjorgensen/Desktop/Underwood 2024/LandUseTables.xlsx - J3.csv")
JMainLandUse<-read.csv("/Users/tjorgensen/Desktop/Underwood 2024/LandUseTables.xlsx - JabezMain.csv")
HBLandUse<-read.csv("/Users/tjorgensen/Desktop/Underwood 2024/LandUseTables.xlsx - HowardsBranch.csv")
ARLandUse<-read.csv("/Users/tjorgensen/Desktop/Underwood 2024/LandUseTables.xlsx - ArthursRun.csv")
CCLandUse<-read.csv("/Users/tjorgensen/Desktop/Underwood 2024/LandUseTables.xlsx - CattailCreek.csv")
CULandUse<-read.csv("/Users/tjorgensen/Desktop/Underwood 2024/LandUseTables.xlsx - CattailCreekUpper.csv")

```

We are interested in three land use metrics: percent pervious surface in
watershed, percent vegetated surface in watershed, and percent tree
cover in wetland. I use the sqldf package to gather and calculate these
(other packages would work but I am most used to SQL language as a GIS
user)

```{r}

J3PerviousPct<-sum(sqldf("select Pct from J3LandUse where Porosity = 'Yes'"))
JMainPerviousPct<-sum(sqldf("SELECT Pct FROM JMainLandUse WHERE Porosity = 'Yes'"))
HBPerviousPct<-sum(sqldf("SELECT Pct FROM HBLandUse WHERE Porosity = 'Yes'"))
ARPerviousPct<-sum(sqldf("SELECT Pct FROM ARLandUse WHERE Porosity = 'Yes'"))
CCPerviousPct<-sum(sqldf("SELECT Pct FROM CCLandUse WHERE Porosity = 'Yes'"))
CUPerviousPct<-sum(sqldf("SELECT Pct FROM CULandUse WHERE Porosity = 'Yes'"))
J3VegPct<-sum(sqldf("SELECT Pct FROM J3LandUse WHERE Vegetated = 'Yes'"))
JMainVegPct<-sum(sqldf("SELECT Pct FROM JMainLandUse WHERE Vegetated = 'Yes'"))
HBVegPct<-sum(sqldf("SELECT Pct FROM HBLandUse WHERE Vegetated = 'Yes'"))
ARVegPct<-sum(sqldf("SELECT Pct FROM ARLandUse WHERE Vegetated = 'Yes'"))
CCVegPct<-sum(sqldf("SELECT Pct FROM CCLandUse WHERE Vegetated = 'Yes'"))
CUVegPct<-sum(sqldf("SELECT Pct FROM CULandUse WHERE Vegetated = 'Yes'"))
J3TreePct<-sum(sqldf("SELECT Pct FROM J3LandUse WHERE TreeCvr = 'Yes'"))
JMainTreePct<-sum(sqldf("SELECT Pct FROM JMainLandUse WHERE TreeCvr = 'Yes'"))
HBTreePct<-sum(sqldf("SELECT Pct FROM HBLandUse WHERE TreeCvr = 'Yes'"))
ARTreePct<-sum(sqldf("SELECT Pct FROM ARLandUse WHERE TreeCvr = 'Yes'"))
CCTreePct<-sum(sqldf("SELECT Pct FROM CCLandUse WHERE TreeCvr = 'Yes'"))
CUTreePct<-sum(sqldf("SELECT Pct FROM CULandUse WHERE TreeCvr = 'Yes'"))

```

Focusing now on manipulating the macroinvertebrates table, I find that
the creation of certan data later on will become easier if we give each
sampling site its own dataframe

```{r}
J31Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'J3' AND SiteNo = '1'"))
J32Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'J3' AND SiteNo = '2'"))
J33Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'J3' AND SiteNo = '3'"))
JMain1Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'J1' AND SiteNo = '1'"))
JMain2Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'J1' AND SiteNo = '2'"))
JMain3Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'J1' AND SiteNo = '3'"))
HB1Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'HB' AND SiteNo = '1'"))
HB2Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'HB' AND SiteNo = '2'"))
HB3Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'HB' AND SiteNo = '3'"))
AR1Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'AR' AND SiteNo = '1'"))
AR2Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'AR' AND SiteNo = '2'"))
AR3Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'AR' AND SiteNo = '3'"))
CC1Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'CC' AND SiteNo = '1'"))
CC2Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'CC' AND SiteNo = '2'"))
CC3Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'CC' AND SiteNo = '3'"))
CU1Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'CU' AND SiteNo = '1'"))
CU2Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'CU' AND SiteNo = '2'"))
CU3Tbl<-data.frame(sqldf("SELECT * FROM MacrosDraft WHERE Site = 'CU' AND SiteNo = '3'"))
```

Create the table that will be used for our statistical calculations- one
where a series of dependent and independent variables are grouped under
the key attribute of sampling site

```{r,echo=FALSE}
SiteName<-c('Jabez 3.1','Jabez 3.2','Jabez 3.3',
            'Jabez Main.1','Jabez Main.2','Jabez Main.3',
            'Howards Branch.1','Howards Branch.2','Howards Branch.3',
            'Arthurs Run.1','Arthurs Run.2','Arthurs Run.3',
            'Cattail Creek Lower.1','Cattail Creek Lower.2','Cattail Creek Lower.3',
            'Cattail Creek Upper.1','Cattail Creek Upper.2','Cattail Creek Upper.3')
Restored<-c(rep(TRUE,times=3),rep(FALSE,times=3),rep(TRUE,times=3),
           rep(FALSE,times=3),rep(TRUE,times=3),rep(FALSE,times=3))
Habitat<-c("R","V","R","U","R","U","R","V","R","L","L","L","R","R","V","U","U","V")
PerviousPct<-c(rep(J3PerviousPct,times=3),rep(JMainPerviousPct,times=3),rep(HBPerviousPct,times=3),
               rep(ARPerviousPct,times=3),rep(CCPerviousPct,times=3),rep(CUPerviousPct,times=3))
VegPct<-c(rep(J3VegPct,times=3),rep(JMainVegPct,times=3),rep(HBVegPct,times=3),
               rep(ARVegPct,times=3),rep(CCVegPct,times=3),rep(CUVegPct,times=3))
TreePct<-c(rep(J3TreePct,times=3),rep(JMainTreePct,times=3),rep(HBTreePct,times=3),
               rep(ARTreePct,times=3),rep(CCTreePct,times=3),rep(CUTreePct,times=3))
Size<-c(rep(3032553,times=3),rep(10389250,times=3),rep(948410,times=3),
        rep(428450,times=3),rep(6373450,times=3),rep(4471850,times=3))
FlowRate<-c(rep(0.00894*1000000,times=3),rep(0.07565*1000000,times=3),rep(0.00533*1000000,times=3),
            rep(0.00294*1000000,times=3),rep(0.04446*1000000,times=3),rep(0.02798*1000000,times=3))
MacroTally<-c(sum(J31Tbl$GraphCount),sum(J32Tbl$GraphCount),sum(J33Tbl$GraphCount),
             sum(JMain1Tbl$GraphCount),sum(JMain2Tbl$GraphCount),sum(JMain3Tbl$GraphCount),
             sum(HB1Tbl$GraphCount),sum(HB2Tbl$GraphCount),sum(HB3Tbl$GraphCount),
             sum(AR1Tbl$GraphCount),sum(AR2Tbl$GraphCount),sum(AR3Tbl$GraphCount),
             sum(CC1Tbl$GraphCount),sum(CC2Tbl$GraphCount),sum(CC3Tbl$GraphCount),
             sum(CU1Tbl$GraphCount),sum(CU2Tbl$GraphCount),sum(CU3Tbl$GraphCount))
TotalMass<-c(sum(J31Tbl$Weight,na.rm = TRUE),sum(J32Tbl$Weight,na.rm = TRUE),sum(J33Tbl$Weight,na.rm = TRUE),
             sum(JMain1Tbl$Weight,na.rm = TRUE),sum(JMain2Tbl$Weight,na.rm = TRUE),sum(JMain3Tbl$Weight,na.rm = TRUE),
             sum(HB1Tbl$Weight,na.rm = TRUE),sum(HB2Tbl$Weight,na.rm = TRUE),sum(HB3Tbl$Weight,na.rm = TRUE),
             sum(AR1Tbl$Weight,na.rm = TRUE),sum(AR2Tbl$Weight,na.rm = TRUE),sum(AR3Tbl$Weight,na.rm = TRUE),
             sum(CC1Tbl$Weight,na.rm = TRUE),sum(CC2Tbl$Weight,na.rm = TRUE),sum(CC3Tbl$Weight,na.rm = TRUE),
             sum(CU1Tbl$Weight,na.rm = TRUE),sum(CU2Tbl$Weight,na.rm = TRUE),sum(CU3Tbl$Weight,na.rm = TRUE))
EPTAsPct<-c((sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J3' AND SiteNo = '1' AND TaxOrder = 'Ephemeroptera' OR Site = 'J3' AND SiteNo = '1' AND TaxOrder = 'Plecoptera' OR Site = 'J3' AND SiteNo = '1' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J3' AND SiteNo = '1' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J3' AND SiteNo = '2' AND TaxOrder = 'Ephemeroptera' OR Site = 'J3' AND SiteNo = '2' AND TaxOrder = 'Plecoptera' OR Site = 'J3' AND SiteNo = '2' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J3' AND SiteNo = '2' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J3' AND SiteNo = '3' AND TaxOrder = 'Ephemeroptera' OR Site = 'J3' AND SiteNo = '3' AND TaxOrder = 'Plecoptera' OR Site = 'J3' AND SiteNo = '3' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J3' AND SiteNo = '3' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J1' AND SiteNo = '1' AND TaxOrder = 'Ephemeroptera' OR Site = 'J1' AND SiteNo = '1' AND TaxOrder = 'Plecoptera' OR Site = 'J1' AND SiteNo = '1' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J1' AND SiteNo = '1' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J1' AND SiteNo = '2' AND TaxOrder = 'Ephemeroptera' OR Site = 'J1' AND SiteNo = '2' AND TaxOrder = 'Plecoptera' OR Site = 'J1' AND SiteNo = '2' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J1' AND SiteNo = '2' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J1' AND SiteNo = '3' AND TaxOrder = 'Ephemeroptera' OR Site = 'J1' AND SiteNo = '3' AND TaxOrder = 'Plecoptera' OR Site = 'J1' AND SiteNo = '3' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'J1' AND SiteNo = '3' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'HB' AND SiteNo = '1' AND TaxOrder = 'Ephemeroptera' OR Site = 'HB' AND SiteNo = '1' AND TaxOrder = 'Plecoptera' OR Site = 'HB' AND SiteNo = '1' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'HB' AND SiteNo = '1' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'HB' AND SiteNo = '2' AND TaxOrder = 'Ephemeroptera' OR Site = 'HB' AND SiteNo = '2' AND TaxOrder = 'Plecoptera' OR Site = 'HB' AND SiteNo = '2' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'HB' AND SiteNo = '2' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'HB' AND SiteNo = '3' AND TaxOrder = 'Ephemeroptera' OR Site = 'HB' AND SiteNo = '3' AND TaxOrder = 'Plecoptera' OR Site = 'HB' AND SiteNo = '3' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'HB' AND SiteNo = '3' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'AR' AND SiteNo = '1' AND TaxOrder = 'Ephemeroptera' OR Site = 'AR' AND SiteNo = '1' AND TaxOrder = 'Plecoptera' OR Site = 'AR' AND SiteNo = '1' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'AR' AND SiteNo = '1' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'AR' AND SiteNo = '2' AND TaxOrder = 'Ephemeroptera' OR Site = 'AR' AND SiteNo = '2' AND TaxOrder = 'Plecoptera' OR Site = 'AR' AND SiteNo = '2' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'AR' AND SiteNo = '2' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'AR' AND SiteNo = '3' AND TaxOrder = 'Ephemeroptera' OR Site = 'AR' AND SiteNo = '3' AND TaxOrder = 'Plecoptera' OR Site = 'AR' AND SiteNo = '3' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'AR' AND SiteNo = '3' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CC' AND SiteNo = '1' AND TaxOrder = 'Ephemeroptera' OR Site = 'CC' AND SiteNo = '1' AND TaxOrder = 'Plecoptera' OR Site = 'CC' AND SiteNo = '1' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CC' AND SiteNo = '1' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CC' AND SiteNo = '2' AND TaxOrder = 'Ephemeroptera' OR Site = 'CC' AND SiteNo = '2' AND TaxOrder = 'Plecoptera' OR Site = 'CC' AND SiteNo = '2' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CC' AND SiteNo = '2' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CC' AND SiteNo = '3' AND TaxOrder = 'Ephemeroptera' OR Site = 'CC' AND SiteNo = '3' AND TaxOrder = 'Plecoptera' OR Site = 'CC' AND SiteNo = '3' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CC' AND SiteNo = '3' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CU' AND SiteNo = '1' AND TaxOrder = 'Ephemeroptera' OR Site = 'CU' AND SiteNo = '1' AND TaxOrder = 'Plecoptera' OR Site = 'CU' AND SiteNo = '1' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CU' AND SiteNo = '1' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CU' AND SiteNo = '2' AND TaxOrder = 'Ephemeroptera' OR Site = 'CU' AND SiteNo = '2' AND TaxOrder = 'Plecoptera' OR Site = 'CU' AND SiteNo = '2' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CU' AND SiteNo = '2' ")))*100,
            (sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CU' AND SiteNo = '3' AND TaxOrder = 'Ephemeroptera' OR Site = 'CU' AND SiteNo = '3' AND TaxOrder = 'Plecoptera' OR Site = 'CU' AND SiteNo = '3' AND TaxOrder = 'Trichoptera'"))/sum(sqldf("SELECT GraphCount FROM MacrosDraft WHERE Site = 'CU' AND SiteNo = '3' ")))*100)

FinalDataFrameHABITAT<-data.frame(SiteName,Restored,Habitat,Size,FlowRate)
FinalDataFrameHABITAT$DO<-SiteData$DO
FinalDataFrameHABITAT$Temperature<-SiteData$Temp
FinalDataFrameHABITAT$Conductivity<-SiteData$Cond
FinalDataFrameHABITAT$PebbleSize<-((SiteData$PblCountL+SiteData$PblCountW+SiteData$PblCountD)/3)
FinalDataFrameHABITAT$PerviousPct<-PerviousPct
FinalDataFrameHABITAT$VegPct<-VegPct
FinalDataFrameHABITAT$TreePct<-TreePct
FinalDataFrameHABITAT$Date<-SiteData$Date
FinalDataFrameHABITAT$MacroTally<-MacroTally
FinalDataFrameHABITAT$TotalMass<-TotalMass
FinalDataFrameHABITAT$EPTAsPct<-EPTAsPct
TaxaCount <- MacrosDraft %>% 
  pivot_longer(Family) %>% 
  count(name, value)
print(TaxaCount)
data.frame(TaxaCount)
J31TaxaCount <- J31Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(J31TaxaCount)
J32TaxaCount <- J32Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(J32TaxaCount)
J33TaxaCount <- J33Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(J33TaxaCount)
JMain1TaxaCount <- JMain1Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(JMain1TaxaCount)
JMain2TaxaCount <- JMain2Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(JMain2TaxaCount)
JMain3TaxaCount <- JMain3Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(JMain3TaxaCount)
HB1TaxaCount <- HB1Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(HB1TaxaCount)
HB2TaxaCount <- HB2Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(HB2TaxaCount)
HB3TaxaCount <- HB3Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(HB3TaxaCount)
AR1TaxaCount <- AR1Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(AR1TaxaCount)
AR2TaxaCount <- AR2Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(AR2TaxaCount)
AR3TaxaCount <- AR3Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(AR3TaxaCount)
CC1TaxaCount <- CC1Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(CC1TaxaCount)
CC2TaxaCount <- CC2Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(CC2TaxaCount)
CC3TaxaCount <- CC3Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(CC3TaxaCount)
CU1TaxaCount <- CU1Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(CU1TaxaCount)
CU2TaxaCount <- CU2Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(CU2TaxaCount)
CU3TaxaCount <- CU3Tbl %>% 
  pivot_longer(Family) %>% 
  count(name, value)
data.frame(CU3TaxaCount)
FinalDataFrameHABITAT$NumTaxa<- c(count(J31TaxaCount),count(J32TaxaCount),count(J33TaxaCount),
                                  count(JMain1TaxaCount),count(JMain2TaxaCount),count(JMain3TaxaCount),
                                  count(HB1TaxaCount),count(HB2TaxaCount),count(HB3TaxaCount),
                                  count(AR1TaxaCount),count(AR2TaxaCount),count(AR3TaxaCount),
                                  count(CC1TaxaCount),count(CC2TaxaCount),count(CC3TaxaCount),
                                  count(CU1TaxaCount),count(CU2TaxaCount),count(CU3TaxaCount))
J31Shannon<- -sum((J31TaxaCount$n/sum(J31TaxaCount$n))*log((J31TaxaCount$n/sum(J31TaxaCount$n))))
J32Shannon<- -sum((J32TaxaCount$n/sum(J32TaxaCount$n))*log((J32TaxaCount$n/sum(J32TaxaCount$n))))
J33Shannon<- -sum((J33TaxaCount$n/sum(J33TaxaCount$n))*log((J33TaxaCount$n/sum(J33TaxaCount$n))))
JMain1Shannon<- -sum((JMain1TaxaCount$n/sum(JMain1TaxaCount$n))*log((JMain1TaxaCount$n/sum(JMain1TaxaCount$n))))
JMain2Shannon<- -sum((JMain2TaxaCount$n/sum(JMain2TaxaCount$n))*log((JMain2TaxaCount$n/sum(JMain2TaxaCount$n))))
JMain3Shannon<- -sum((JMain3TaxaCount$n/sum(JMain3TaxaCount$n))*log((JMain3TaxaCount$n/sum(JMain3TaxaCount$n))))
HB1Shannon<- -sum((HB1TaxaCount$n/sum(HB1TaxaCount$n))*log((HB1TaxaCount$n/sum(HB1TaxaCount$n))))
HB2Shannon<- -sum((HB2TaxaCount$n/sum(HB2TaxaCount$n))*log((HB2TaxaCount$n/sum(HB2TaxaCount$n))))
HB3Shannon<- -sum((HB3TaxaCount$n/sum(HB3TaxaCount$n))*log((HB3TaxaCount$n/sum(HB3TaxaCount$n))))
AR1Shannon<- -sum((AR1TaxaCount$n/sum(AR1TaxaCount$n))*log((AR1TaxaCount$n/sum(AR1TaxaCount$n))))
AR2Shannon<- -sum((AR2TaxaCount$n/sum(AR2TaxaCount$n))*log((AR2TaxaCount$n/sum(AR2TaxaCount$n))))
AR3Shannon<- -sum((AR3TaxaCount$n/sum(AR3TaxaCount$n))*log((AR3TaxaCount$n/sum(AR3TaxaCount$n))))
CC1Shannon<- -sum((CC1TaxaCount$n/sum(CC1TaxaCount$n))*log((CC1TaxaCount$n/sum(CC1TaxaCount$n))))
CC2Shannon<- -sum((CC2TaxaCount$n/sum(CC2TaxaCount$n))*log((CC2TaxaCount$n/sum(CC2TaxaCount$n))))
CC3Shannon<- -sum((CC3TaxaCount$n/sum(CC3TaxaCount$n))*log((CC3TaxaCount$n/sum(CC3TaxaCount$n))))
CU1Shannon<- -sum((CU1TaxaCount$n/sum(CU1TaxaCount$n))*log((CU1TaxaCount$n/sum(CU1TaxaCount$n))))
CU2Shannon<- -sum((CU2TaxaCount$n/sum(CU2TaxaCount$n))*log((CU2TaxaCount$n/sum(CU2TaxaCount$n))))
CU3Shannon<- -sum((CU3TaxaCount$n/sum(CU3TaxaCount$n))*log((CU3TaxaCount$n/sum(CU3TaxaCount$n))))
FinalDataFrameHABITAT$ShannonDiversity<- c(J31Shannon,J32Shannon,J33Shannon,JMain1Shannon,JMain2Shannon,JMain3Shannon,HB1Shannon,HB2Shannon,HB3Shannon,AR1Shannon,AR2Shannon,AR3Shannon,CC1Shannon,CC2Shannon,CC3Shannon,CU1Shannon,CU2Shannon,CU3Shannon)
FFGCount <- MacrosDraft %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(FFGCount)
J31FFGCount <- J31Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(J31FFGCount)
J32FFGCount <- J32Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(J32FFGCount)
J33FFGCount <- J33Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(J33FFGCount)
JMain1FFGCount <- JMain1Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(JMain1FFGCount)
JMain2FFGCount <- JMain2Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(JMain2FFGCount)
JMain3FFGCount <- JMain3Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(JMain3FFGCount)
HB1FFGCount <- HB1Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(HB1FFGCount)
HB2FFGCount <- HB2Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(HB2FFGCount)
HB3FFGCount <- HB3Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(HB3FFGCount)
AR1FFGCount <- AR1Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(AR1FFGCount)
AR2FFGCount <- AR2Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(AR2FFGCount)
AR3FFGCount <- AR3Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(AR3FFGCount)
CC1FFGCount <- CC1Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(CC1FFGCount)
CC2FFGCount <- CC2Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(CC2FFGCount)
CC3FFGCount <- CC3Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(CC3FFGCount)
CU1FFGCount <- CU1Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(CU1FFGCount)
CU2FFGCount <- CU2Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(CU2FFGCount)
CU3FFGCount <- CU3Tbl %>% 
  pivot_longer(FFG) %>% 
  count(name, value)
data.frame(CU3FFGCount)
FinalDataFrameHABITAT$NumFFG<- c(count(J31FFGCount),count(J32FFGCount),count(J33FFGCount),
                                  count(JMain1FFGCount),count(JMain2FFGCount),count(JMain3FFGCount),
                                  count(HB1FFGCount),count(HB2FFGCount),count(HB3FFGCount),
                                  count(AR1FFGCount),count(AR2FFGCount),count(AR3FFGCount),
                                  count(CC1FFGCount),count(CC2FFGCount),count(CC3FFGCount),
                                  count(CU1FFGCount),count(CU2FFGCount),count(CU3FFGCount))
J31FFGShannon<- -sum((J31FFGCount$n/sum(J31FFGCount$n))*log((J31FFGCount$n/sum(J31FFGCount$n))))
J32FFGShannon<- -sum((J32FFGCount$n/sum(J32FFGCount$n))*log((J32FFGCount$n/sum(J32FFGCount$n))))
J33FFGShannon<- -sum((J33FFGCount$n/sum(J33FFGCount$n))*log((J33FFGCount$n/sum(J33FFGCount$n))))
JMain1FFGShannon<- -sum((JMain1FFGCount$n/sum(JMain1FFGCount$n))*log((JMain1FFGCount$n/sum(JMain1FFGCount$n))))
JMain2FFGShannon<- -sum((JMain2FFGCount$n/sum(JMain2FFGCount$n))*log((JMain2FFGCount$n/sum(JMain2FFGCount$n))))
JMain3FFGShannon<- -sum((JMain3FFGCount$n/sum(JMain3FFGCount$n))*log((JMain3FFGCount$n/sum(JMain3FFGCount$n))))
HB1FFGShannon<- -sum((HB1FFGCount$n/sum(HB1FFGCount$n))*log((HB1FFGCount$n/sum(HB1FFGCount$n))))
HB2FFGShannon<- -sum((HB2FFGCount$n/sum(HB2FFGCount$n))*log((HB2FFGCount$n/sum(HB2FFGCount$n))))
HB3FFGShannon<- -sum((HB3FFGCount$n/sum(HB3FFGCount$n))*log((HB3FFGCount$n/sum(HB3FFGCount$n))))
AR1FFGShannon<- -sum((AR1FFGCount$n/sum(AR1FFGCount$n))*log((AR1FFGCount$n/sum(AR1FFGCount$n))))
AR2FFGShannon<- -sum((AR2FFGCount$n/sum(AR2FFGCount$n))*log((AR2FFGCount$n/sum(AR2FFGCount$n))))
AR3FFGShannon<- -sum((AR3FFGCount$n/sum(AR3FFGCount$n))*log((AR3FFGCount$n/sum(AR3FFGCount$n))))
CC1FFGShannon<- -sum((CC1FFGCount$n/sum(CU1FFGCount$n))*log((CC1FFGCount$n/sum(CC1FFGCount$n))))
CC2FFGShannon<- -sum((CC2FFGCount$n/sum(CU2FFGCount$n))*log((CC2FFGCount$n/sum(CC2FFGCount$n))))
CC3FFGShannon<- -sum((CC3FFGCount$n/sum(CU3FFGCount$n))*log((CC3FFGCount$n/sum(CC3FFGCount$n))))
CU1FFGShannon<- -sum((CU1FFGCount$n/sum(CU1FFGCount$n))*log((CU1FFGCount$n/sum(CU1FFGCount$n))))
CU2FFGShannon<- -sum((CU2FFGCount$n/sum(CU2FFGCount$n))*log((CU2FFGCount$n/sum(CU2FFGCount$n))))
CU3FFGShannon<- -sum((CU3FFGCount$n/sum(CU3FFGCount$n))*log((CU3FFGCount$n/sum(CU3FFGCount$n))))
FinalDataFrameHABITAT$FFGDiversity<- c(J31FFGShannon,J32FFGShannon,J33FFGShannon,JMain1FFGShannon,JMain2FFGShannon,JMain3FFGShannon,HB1FFGShannon,HB2FFGShannon,HB3FFGShannon,AR1FFGShannon,AR2FFGShannon,AR3FFGShannon,CC1FFGShannon,CC2FFGShannon,CC3FFGShannon,CU1FFGShannon,CU2FFGShannon,CU3FFGShannon)

```

INDEPENDENT VARIABLES SiteName= the name of the river the sample was
taken from plus a 1 2 or 3 indicating the downstream to upstream-most
position Restored= whether or not the site had undergone RSC engineering
by Underwood Habitat= qualitative measurement of what might attract
macros to the area R: Cobble riffle V: Aquatic vegetation (living) U:
Shaded undercut bank L: Dead layers of leaf litter Size= Square meterage
of the watershed, measured from downstream-most site FlowRate= Cubic
centimeters per second of water moving through a carefully-chosen
transect of the given river DO= Dissolved oxygen measured in miligrams
per liter by YSI sampler Temperature= Temperature of water in degrees
Celsius measured by YSI sampler Conductivity= Conductivity of water in
S/cm measured by YSI sampler PblCount= Average of the three longest
pebble axes, taken for 25 pebbles at each site (modified Wolman pebble
count) PerviousPct= % of river's watershed that is covered in a pervious
surface (forest cover, cropland, etc) VegPct= % of river's watershed
that is actively covered in vegetation TreePct= % of river's watershed
that is wooded Date= day of year sample was taken

DEPENDENT VARIABLES MacroTally= Raw number of macroinvertebrates at each
site TotalMass= Sum of mass (in mg) of all macroinvertebrates at
sampling site EPTAsPct= The % of the total number of macroinvertebrates
that were Ephemeroptera (mayflies), Plecoptera (stoneflies), or
Trichoptera (caddisflies). These three orders are of special importance
because they are highly disturbance-sensitive. NumTaxa= Number of
distinct taxa at a sampling site ShannonDiversity= Site's Shannon
Diversity Index value, a good metric that factors in both species
richness and species evenness NumFFG= Number of distinct functional
feeding groups (shredder, scraper, gatherer, predator, filterer)
represented in a sampling site's biota FFGDiversity= The Shannon
Diversity Index but run on FFG instead of taxa

##PART TWO ANSWERING THE QUESTIONS ~ STATISTICAL TESTS
Question 1- Are restored stream-wetland complexes supporting a higher
biomass and functional feeding group diversity of macroinvertebrates
than unrestored streams? 
Independent variable: Restored
Dependent variable: TotalMass, EPTAsPct, NumFFG, FFGDiversity (Qualitative vs. Quantitative- best to use Kruskal-Wallis to test and bar graphs to display)

```{r}
kruskal.test(MacroTally~Restored,data=FinalDataFrameHABITAT)

kruskal.test(TotalMass~Restored,data=FinalDataFrameHABITAT)

kruskal.test(EPTAsPct~Restored,data=FinalDataFrameHABITAT)

kruskal.test(FFGDiversity~Restored,data=FinalDataFrameHABITAT)

kruskal.test(ShannonDiversity~Restored,data=FinalDataFrameHABITAT)

```
Also of interest: are any of these strongly correlated with the type of habitat (also a quantitative measure)?

```{r}
kruskal.test(MacroTally~Habitat,data=FinalDataFrameHABITAT)

kruskal.test(TotalMass~Habitat,data=FinalDataFrameHABITAT)

kruskal.test(EPTAsPct~Habitat,data=FinalDataFrameHABITAT)

kruskal.test(FFGDiversity~Habitat,data=FinalDataFrameHABITAT)

kruskal.test(ShannonDiversity~Habitat,data=FinalDataFrameHABITAT)

```
Total biomass per site is statistically correlated with both restoration status and habitat type- in other words there is a significant difference in these values between restored and unrestored sites. We can use the pairwise wilcox test to determine which of these groups were highest.

```{r}
pairwise.wilcox.test(FinalDataFrameHABITAT$TotalMass, FinalDataFrameHABITAT$Restored,
                 p.adjust.method = "BH")
pairwise.wilcox.test(FinalDataFrameHABITAT$TotalMass, FinalDataFrameHABITAT$Habitat,
                 p.adjust.method = "BH")

```

Question 2- What is the relationship between macroinvertebrate community health and the surrounding land use of the area?
Independent Variable: Land Use variables
Dependent variable: MacroTally (Spearman rank correlation)

```{r}
cor.test(x=FinalDataFrameHABITAT$TreePct,y=FinalDataFrameHABITAT$MacroTally,method='spearman')
cor.test(x=FinalDataFrameHABITAT$VegPct,y=FinalDataFrameHABITAT$MacroTally,method='spearman')
cor.test(x=FinalDataFrameHABITAT$PerviousPct,y=FinalDataFrameHABITAT$MacroTally,method='spearman')

ggplot(data=FinalDataFrameHABITAT,mapping=aes(x=TreePct,y=MacroTally))+geom_point()

```

Question 3- Does the biomass and functional feeding group diversity of macroinvertebrates in restored streams depend on the surrounding land use?
Independent: Land Use variables
Dependent: TotalMass, EPTAsPct, NumFFG, FFGDiversity (Quantitative vs. Quantitative- needs Spearman rank correlation

```{r}
cor.test(x=FinalDataFrameHABITAT$TreePct,y=FinalDataFrameHABITAT$TotalMass,method='spearman')
cor.test(x=FinalDataFrameHABITAT$TreePct,y=FinalDataFrameHABITAT$EPTAsPct,method='spearman')
cor.test(x=FinalDataFrameHABITAT$TreePct,y=FinalDataFrameHABITAT$ShannonDiversity,method='spearman')
cor.test(x=FinalDataFrameHABITAT$TreePct,y=FinalDataFrameHABITAT$FFGDiversity,method='spearman')
cor.test(x=FinalDataFrameHABITAT$VegPct,y=FinalDataFrameHABITAT$TotalMass,method='spearman')
cor.test(x=FinalDataFrameHABITAT$VegPct,y=FinalDataFrameHABITAT$EPTAsPct,method='spearman')
cor.test(x=FinalDataFrameHABITAT$VegPct,y=FinalDataFrameHABITAT$ShannonDiversity,method='spearman')
cor.test(x=FinalDataFrameHABITAT$VegPct,y=FinalDataFrameHABITAT$FFGDiversity,method='spearman')
cor.test(x=FinalDataFrameHABITAT$PerviousPct,y=FinalDataFrameHABITAT$TotalMass,method='spearman')
cor.test(x=FinalDataFrameHABITAT$PerviousPct,y=FinalDataFrameHABITAT$EPTAsPct,method='spearman')
cor.test(x=FinalDataFrameHABITAT$PerviousPct,y=FinalDataFrameHABITAT$ShannonDiversity,method='spearman')
cor.test(x=FinalDataFrameHABITAT$PerviousPct,y=FinalDataFrameHABITAT$FFGDiversity,method='spearman')

```

